MS Prelim Data K23

### 2024-06-07 scripts/first_pass_adapting_K23_analysis_20231031.Rmd
adding all sorts of things for within/bet network
After using neurosynth maps

```{r add_suffix_to_each_measure}

@@

#read in volumes of fascicles and make columns reflecting volume of disease injured per fascicle

@@

@@

#summarize means across all fascicles
```{r make_summary_measures_of_fascicle_burden}

#n=812
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  mutate(total_fascicle_prop_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_prop_suffix)))) %>%
  mutate(total_fascicle_volume_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_vol_suffix))))
```

#Compare disease burden by anxiety comorbidity
```{r compare_disease_burden_by_anxiety_comorbidity}

#lm
fascicle_number_of_anxiety_dx_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ number_of_anxiety_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_anxiety_dx_lm) <- fascicle_names_w_prop_suffix
fascicle_number_of_anxiety_dx_anova <- lapply(fascicle_number_of_anxiety_dx_lm, anova)

@@

```
#Has depression AND anxiety diagnoses
```{r compare_disease_burden_by_dep_AND_anxiety_comorbidity}

@@

@@

anxiety_vs_no_psych <- df_demo_and_fascicles %>%
  mutate(anxietyGroupVar = ifelse((Has.anxietydx | On.Anxiolytics_no_beta_blocker), 2, ifelse(healthy_ish, 1, 0))) %>% #complicated, if they have an anxiety disorder, they get a 2, if they are health-ish, no psych dx or meds, 1, otherwise, exclude
  filter(anxietyGroupVar != 0) %>% #get rid of the people who are not anxious and not healthy_ish, n =643, lost n=169
  filter(is.na(depGroupVar)) %>% #exclude people with depression (87 have both depression and anxiety), n = 557, anxiety = 101, nonanxious = 456)
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup

@@

>>>>

## Read in lesion volume size
```{r read_in_lesion_volumes}
##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv("/Users/eballer/BBL/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes", sep = ",", header = TRUE) #n=958 unique, identical to fascicle unique
lesion_volumes <- lesion_volumes %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

## Merge with data frame with good mimosas
```{r merge_lesion_volume_w_mimosa_df}
##########################################################
#             Merge lesion volumes with mimosa df        #
##########################################################
#merge whole data set with fascicle proportions, n = 812
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  left_join(lesion_volumes, by = c("EMPI", "EXAM_DATE"))

## Make separate columns for proportional injury (prop), total volume injury (vol)
```{r add_suffix_to_each_measure}

########################################
## Read in Fiber Volume Values        ##
########################################

@@

@@

```
##Anxiety Network from neurosynth (association, not uniformity)
###read in file with fasciclexanxiety overlap, and define anxiety network
```{r read_in_neurosynth_anxiety_network}
neurosynth_anxiety_network_fascicle_overlap <- read.csv("/Users/eballer/BBL/msanxiety/results/streamline_volume_within_anxiety_network_association_neuro_maps_FDR_01.csv", header = T, sep = ",") %>% 
  filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves))) #get rid of cranial nerves up front

fascicles_with_any_anxiety_overlap <- neurosynth_anxiety_network_fascicle_overlap %>%
  filter(!is.na(prop_in_mask))  #n = 43

#top 25% 
number_of_brain_regions_in_top_25 <- round((dim(fascicles_with_any_anxiety_overlap)[1]/4)) #n=11

fascicles_with_vol_top_25_percent_anxiety_overlap <- neurosynth_anxiety_network_fascicle_overlap %>%
  filter(!is.na(prop_in_mask)) %>%
  arrange(desc(non_zero_voxels_in_anxiety_map)) %>%
  slice(1:number_of_brain_regions_in_top_25)

fascicles_with_prop_top_25_percent_anxiety_overlap <- neurosynth_anxiety_network_fascicle_overlap %>%
  filter(!is.na(prop_in_mask)) %>%
  arrange(desc(prop_in_mask)) %>%
  slice(1:number_of_brain_regions_in_top_25)

#name vectors to use later
fascicle_names_any_anxiety_overlap_vol <- fascicles_with_any_anxiety_overlap$fascicle 
fascicle_names_top_25_vol <- fascicles_with_vol_top_25_percent_anxiety_overlap$fascicle 
fascicle_names_bottom_75_vol <- fascicles_with_any_anxiety_overlap %>% filter(!(fascicle %in% fascicle_names_top_25_vol)) %>% pull(fascicle)

fascicle_names_any_anxiety_overlap_prop <- fascicles_with_any_anxiety_overlap$fascicle
fascicle_names_top_25_prop <-fascicles_with_prop_top_25_percent_anxiety_overlap$fascicle 
fascicle_names_bottom_75_prop <- fascicles_with_any_anxiety_overlap %>% filter(!(fascicle %in% fascicle_names_top_25_prop)) %>% pull(fascicle)

##add vol/prop
fascicle_names_any_anxiety_overlap_vol_w_suff <- fascicle_names_any_anxiety_overlap_vol %>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
fascicle_names_top_25_vol_w_suff <- fascicle_names_top_25_vol %>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
fascicle_names_bottom_75_vol_w_suff <- fascicle_names_bottom_75_vol  %>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #take fascicles not in

fascicle_names_any_anxiety_overlap_prop_w_suff <- fascicle_names_any_anxiety_overlap_prop %>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_top_25_prop_w_suff <- fascicle_names_top_25_prop %>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_bottom_75_prop_w_suff <- fascicle_names_bottom_75_prop %>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)

# Total volume of injury within each anxiety network, double checked that : top_25_total_volume + bottom_75_total_volume
total_fascicle_volume <- sum(neurosynth_anxiety_network_fascicle_overlap$num_voxels_total)
any_anxiety_overlap_total_volume <- sum(fascicles_with_any_anxiety_overlap$num_voxels_total) # vol = 1093606
top_25_total_volume <-sum(fascicles_with_vol_top_25_percent_anxiety_overlap$num_voxels_total) # vol = 548305
bottom_75_total_volume <- sum(fascicles_with_any_anxiety_overlap$num_voxels_total[fascicles_with_any_anxiety_overlap$fascicle %in% fascicle_names_bottom_75_vol]) #vol = 548305

@@

###Add columns in the main data frame indicating overall injury in the whole network, in the top 25 network (vol and prop), and bottom 75%
```{r add_anxiety_network_details_to_data_frame}

#all of this has been checked - making sure prop * vol network = sum of damage

df_demo_and_fascicles <- df_demo_and_fascicles %>%

  mutate(mean_fascicle_prop_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_prop_suffix)))) %>%
  mutate(mean_fascicle_volume_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%

  #sum of total volume lost and proportion of volume lost 
  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_fascicle_volume) %>%

  #sum of total volume lost in anxiety network and proportion of volume lost in anxietyn network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_inanxietynet_all = 
           rowSums(across(all_of(fascicle_names_any_anxiety_overlap_vol_w_suff)))) %>%
  mutate(proportion_volume_lost_anxiety_net_by_network_size_sum = sum_fascicle_vol_lost_inanxietynet_all/any_anxiety_overlap_total_volume) %>%

  #sum of total volume lost in anxiety network and proportion of volume lost in anxiety network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_inanxietynet_top_quartile = 
           rowSums(across(all_of(fascicle_names_top_25_vol_w_suff)))) %>%
  mutate(proportion_volume_lost_anxiety_net_by_network_size_sum_top_25 = sum_fascicle_vol_lost_inanxietynet_top_quartile/top_25_total_volume) %>%

  #sum of total volume lost in nonanxiety network and proportion of volume lost in anxiety network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_nonanxiety_net_bottom_3_quartiles = 
           rowSums(across(all_of(fascicle_names_bottom_75_vol_w_suff)))) %>%
  mutate(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 = sum_fascicle_vol_lost_nonanxiety_net_bottom_3_quartiles/bottom_75_total_volume) %>%

  ungroup()

#neurosynth_anxiety_network_fascicle_overlap
```

## Anxiety within/outside network interaction
```{r anxiety_net_intx}

#clean up data frame
df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(anxietyGroupVar != 0) %>% #lose 93, n now 719
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(anxietyGroupVar==1, "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Anxiety_network = proportion_volume_lost_anxiety_net_by_network_size_sum_top_25) %>%
  mutate(Nonanxiety_network = proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75) %>%
  dplyr::select(subject, osex, Diagnosis, Anxiety_network, Nonanxiety_network, PAT_AGE_AT_EXAM) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#simple t test, disease within versus outside anxiety network; mean within anxiety net: 0.5426694 affected, mean outside anxiety network = 0.3062429, T=57.44, p<2.2*10^-16, 95% CI 0.22-0.24

hist(df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Network == "Anxiety_network"])
hist(df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Network == "Nonanxiety_network"])

within_vs_outside_anxiety_net_t <- t.test(df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Network == "Anxiety_network"], df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Network == "Nonanxiety_network"], paired = T)
summary(within_vs_outside_anxiety_net_t)

 #simple lmer, confirms T
disease_within_vs_outside_anxiety_net_lmer <- lmer(proportion_affected ~ Network + (1|subject), data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_lmer)

#diagnosis: Anxiety > nonanxiety T = 2.4, p = 0.015; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
summary(diagnosis_by_network_interaction_lmer)

#Diagnosis: MS+ > MS- 0.011-0.1; Network dep>nondep -.104--.076; intx: -0.040--0.003
confint(diagnosis_by_network_interaction_lmer)

```
# Compare anxiety burden (presence of condition, number of anxiety comorbidities) by fascicle, by overall injury, and by volume of lesions
```{r compare_anxiety_burden_by_fascicle_injury_lesion_volume}

## Compare disease burden by anxiety comorbidity by fascicle
```{r compare_disease_burden_by_anxiety_comorbidity}

#### in each fascicle
#lm
fascicle_number_of_anxiety_dx_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ number_of_anxiety_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_anxiety_dx_lm) <- fascicle_names_w_prop_suffix
fascicle_number_of_anxiety_dx_anova <- lapply(fascicle_number_of_anxiety_dx_lm, anova)

@@

## Has depression AND anxiety diagnoses
```{r compare_disease_burden_by_dep_AND_anxiety_comorbidity}

@@

### total lesion burden

### lesion volume

@@

anxiety_vs_no_psych <- df_demo_and_fascicles %>%
  mutate(anxietyGroupVar = ifelse((Has.anxietydx | On.Anxiolytics_no_beta_blocker), 2, ifelse(healthy_ish, 1, 0))) %>% #complicated, if they have an anxiety disorder, they get a 2, if they are health-ish, no psych dx or meds, 1, otherwise, exclude; n anxiety = 271, healthyish = 448, unclassified = 93
  mutate(anxietyGroupVar_ordered = ordered(anxietyGroupVar,levels = c(0,1,2), labels = c("unclassified", "healthyish", "anxiety"))) %>%
  filter(anxietyGroupVar_ordered != "unclassified") %>% #get rid of the people who are not anxious and not healthy_ish, n =541, lost n=93
  filter(is.na(depGroupVar)) %>% #exclude people with depression (87 have both depression and anxiety), n = 557, anxiety = 101, nonanxious = 456)
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup

@@

## Compare total disease burden and lesion volume by anxiety comorbidity
```{r total_burden_and_lesion_volume_by_anxiety_comorbdity}

total_burden_by_anxietydx_lm <- lm(anxietyGroupVar~total_fascicle_volume_lost_no_cranial_nerves, data = anxiety_vs_no_psych)
print(summary(total_burden_by_anxietydx_lm))

total_burden_by_anxiety_lm <- lm(number_of_anxiety_comorbidities_extended~total_fascicle_volume_lost_no_cranial_nerves, data = anxiety_vs_no_psych)
print(summary(total_burden_by_anxiety_lm))

total_lesion_volume_by_anxietydx_lm <- lm(anxietyGroupVar ~volume_of_mimosa_lesions, data = anxiety_vs_no_psych)
print(summary(total_lesion_volume_by_anxietydx_lm))

total_lesion_volume_by_anxiety_lm <- lm(number_of_anxiety_comorbidities_extended ~volume_of_mimosa_lesions, data = anxiety_vs_no_psych)
print(summary(total_lesion_volume_by_anxiety_lm))
